\textbf{ Accurate spectrophotometric estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-31 paper/sp.tex
notes to self

% To-do items
% -----------
% - draft introduction
% - draft data+results section
% - draft discussion
% - make sure we address the cross-validation point
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - address the non-convexity of the model (see twitter convo with @andy_l_jones).

@@

\begin{abstract}\noindent
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which geometric parallaxes are not precisely measured by \gaia.
Here we use the \apogee--\gaia\ overlap to train a purely linear model of
continuum-normalized \apogee\ spectroscopy
(plus \gaia, \zmass, and \wise\ photometry)
that predicts absolute magnitude (logarithmic luminosity) for red stars with $0<\logg<2.2$.
The training makes use of a justified likelihood function for the \gaia\ parallaxes,
so that the training can be performed with full use of uncertainties and
without any cuts on parallax or parallax signal-to-noise ratio.
The model includes an L1 regularization that zeros out the contributions of
most spectral pixels to the parallax estimates.
The training is performed
with leave-out subsamples such that no star's astrometry is used even indirectly in its
spectroscopic parallax estimate.
The model is flexible enough---and has enough photometric information---to
correct for extinction by dust without any
input of explicit extinction or reddening estimates.
We re-label each star in the sample
with a new spectroscopic parallax; this new parallax has an empirical (1-sigma)
uncertainty of 9~percent, which is more precise than the \gaia\ parallax
for the vast majority of targets.
Validation with globular and open clusters shows that the spectroscopic parallaxes
are both accurate and precise.
These distance and luminosity estimates open up great opportunities for
mapping the Milky Way disk with the next
generation of spectroscopic surveys, and especially \sdssv.
\end{abstract}

@@

The history and value (and disadvantages)
of data-driven models for these kinds of problems.

@@

(In addition: what did we assume about the \apogee\ data: Good $\logg$ values.

What is this useful for and how do you use them safely?

@@

Come back to SDSS-V and show the rotation curve of the Galaxy.

>>>>

% To-do items
% -----------
% - draft introduction
% - draft data+results section
% - draft discussion
% - make sure we address the cross-validation point
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - address the non-convexity of the model (see twitter convo with @andy_l_jones).
% - Coryn's point that we should carefully explain how our model is geometric.

@@

\begin{abstract}\noindent
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which geometric parallaxes are not precisely measured by \gaia.
Here we use the \apogee--\gaia\ overlap to train a purely linear model of
continuum-normalized \apogee\ spectroscopy
(plus \gaia, \zmass, and \wise\ photometry)
that predicts absolute magnitude (logarithmic luminosity) for red stars with $0<\logg<2.2$.
The training makes use of a justified likelihood function for the \gaia\ parallaxes,
so that the training can be performed with full use of uncertainties and
without any cuts on parallax or parallax signal-to-noise ratio.
There are no physical inputs to the model; it is a pure linear regression designed
to accurately predict astrometric parallaxes.
The model includes an L1 regularization that zeros out the contributions of
most spectral pixels to the parallax estimates.
The training is performed
with leave-out subsamples such that no star's astrometry is used even indirectly in its
spectroscopic parallax estimate.
The model is flexible enough---and has enough photometric information---to
correct for extinction by dust without any
input of explicit extinction or reddening estimates.
We re-label each star in the sample
with a new spectroscopic parallax; this new parallax has an empirical (1-sigma)
uncertainty of 9~percent, which is more precise than the \gaia\ parallax
for the vast majority of targets.
Validation with globular and open clusters shows that the spectroscopic parallaxes
are both accurate and precise.
These distance and luminosity estimates open up great opportunities for
mapping the Milky Way disk with the next
generation of spectroscopic surveys, and especially \sdssv.
\end{abstract}

@@

The history and value (and disadvantages)
of data-driven models for these kinds of problems.
How could it be that we are completely independent of stellar models?

@@

(In addition: what did we assume about the \apogee\ data: Good $\logg$ values.
Okay pseudo-continuum normalization!)

@@

What is this useful for and how do you use them safely?
HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
x and y, colored by metallicity.
We see the kinematic center of the Galaxy! We see the rotation curve.
We see the chemical gradients.

How do you build a proper probabilistic model using these? How is the answer to that
related to the answer for \gaia\ itself?
